---
title: "Getting started with twelvedataR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with twelvedataR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
# The examples need an API key and an internet connection.
# When no key is available (e.g. on CRAN or on a CI run without the
# secret) the code is shown but not executed, so the vignette still builds.
has_key <- nzchar(Sys.getenv("TWELVEDATA_API_KEY"))
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = has_key,
  fig.width = 7,
  fig.height = 4
)
```

```{r, echo = FALSE, results = "asis", eval = !has_key}
cat("> **Note:** no `TWELVEDATA_API_KEY` was found when this vignette was",
    "built, so the code below is shown without output.\n")
```

## Introduction

[Twelve Data](https://twelvedata.com) is a financial market data provider
with a REST API that covers stocks, ETFs, forex and cryptocurrencies.
**twelvedataR** is a small R interface to this API. It handles the HTTP
requests, checks your inputs before anything is sent, turns the JSON
answers into ordinary data frames and gives clear error messages.

The package exports two functions:

| Function            | What it returns                                  |
|---------------------|--------------------------------------------------|
| `get_time_series()` | Historical prices (open, high, low, close, volume) |
| `get_quote()`       | The latest quote for one symbol                   |

## Installation

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Tvths/twelvedataR")
```

## Setting up your API key

1. Create a free account at <https://twelvedata.com> and copy your API key.
2. Open your `.Renviron` file:

```{r, eval = FALSE}
usethis::edit_r_environ()
```

3. Add the line below, save the file and restart R:

```
TWELVEDATA_API_KEY=your_key_here
```

The package reads the key with `Sys.getenv()`, so it is never written in
your scripts and never ends up on GitHub. If the key is missing, every
function stops with a message telling you how to fix it.

**Rate limits.** The free plan allows about 8 requests per minute and 800
per day. Each call to `get_time_series()` or `get_quote()` uses one
request. If you hit the limit (HTTP 429) the package waits and retries
automatically up to three times.

```{r setup, eval = TRUE}
library(twelvedataR)
```

## Historical prices with `get_time_series()`

The most common use is to download daily prices for a stock:

```{r}
aapl <- get_time_series("AAPL", interval = "1day", outputsize = 100)
head(aapl)
str(aapl)
```

The result is a plain `data.frame`, sorted from oldest to newest, with
numeric price columns and a `datetime` column in UTC. That makes it easy to
use directly with base R or any other package.

### Plotting the closing price

```{r}
plot(aapl$datetime, aapl$close, type = "l",
     main = "Apple (AAPL) - daily closing price",
     xlab = "", ylab = "Price (USD)")
```

### Choosing the interval

`interval` controls the time between observations. Valid values are
`"1min"`, `"5min"`, `"15min"`, `"30min"`, `"45min"`, `"1h"`, `"2h"`,
`"4h"`, `"1day"`, `"1week"` and `"1month"`.

```{r}
monthly <- get_time_series("MSFT", interval = "1month", outputsize = 24)
tail(monthly, 3)
```

### A fixed period

Use `start_date` and `end_date` (as `"YYYY-MM-DD"` strings or `Date`
objects) to get a specific period:

```{r}
jan <- get_time_series("AAPL", interval = "1day",
                       start_date = "2024-01-01", end_date = "2024-01-31")
nrow(jan)
range(jan$datetime)
```

### Other asset classes

The same function works for currency pairs and cryptocurrencies. Note that
these have no `volume` column.

```{r}
eurusd <- get_time_series("EUR/USD", interval = "1day", outputsize = 60)
head(eurusd, 3)
```

## A small analysis: daily returns

Because the result is an ordinary data frame, analysis is straightforward.
Here we compute daily returns and summarise them:

```{r}
aapl$return <- c(NA, diff(aapl$close) / head(aapl$close, -1))

summary(aapl$return)
cat("Annualised volatility:",
    round(sd(aapl$return, na.rm = TRUE) * sqrt(252) * 100, 1), "%\n")

hist(aapl$return, breaks = 30,
     main = "Distribution of daily returns (AAPL)", xlab = "Daily return")
```

## The latest quote with `get_quote()`

```{r}
q <- get_quote("AAPL")
q[, c("symbol", "name", "close", "change", "percent_change")]
```

## Error handling

Inputs are checked **before** any request is sent, which saves requests
from your daily quota:

```{r, error = TRUE, eval = TRUE}
get_time_series("AAPL", interval = "3days")
get_time_series("AAPL", outputsize = 10000)
get_time_series(c("AAPL", "MSFT"))
```

If the input is valid but the API cannot answer (for example an unknown
symbol), the error message from Twelve Data is passed on:

```{r, error = TRUE}
get_time_series("NOTAREALTICKER123")
```

## Summary

* Store your key in `.Renviron` as `TWELVEDATA_API_KEY`.
* `get_time_series()` returns historical prices as a sorted data frame.
* `get_quote()` returns the latest quote as a one-row data frame.
* Invalid inputs are caught locally; API errors are reported clearly.
